A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-06-19 paper-hierarchical.tex
changes xianbo

To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical. For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis. Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of 900 $\times$ 390~$\mu$m$^2$. We get N.A. values of 6.2 $\times$ 21.6~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmission coefficient is the ratio of the accepted N.A. over the beam divergence, resulting in 9.5\% (H) and 0.18\% (V)for ESRF-1 and 16.3\% and 0.18\% for EBS. The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF-1 and 2.95\% for EBS, therefore $3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS. 
%\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}

>>>>

To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical. For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis. Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of  390 $\times$ 900 (H $\times$ V) $\mu$m$^2$. 
We get N.A. values of 2.7 $\times$ 4.9~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmission coefficient is the ratio of the accepted N.A. over the beam divergence, resulting in 
4.1\% (H) and 40\% (V)for ESRF-1 and 7.1\% and 40\% for EBS. The overall transmission considers the product of H and V ratios, therefore 
0.21\% for ESRF-1 and 2.85\% for EBS, therefore 
$4\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS. 
%\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}
